---
title: "Clean and transform: Australian energy report 2021"
author: "CW Yong"
date: "September 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
```

## Project brief
The high level goal of this project is to answer two question's about Australia's energy:

1) From what sources does it come?
2) What role is played by renewable energy?

The raw data used for this analysis comes from the website of the Australian Government's Department of Industry, Science, Energy and Resources. A copy of the Excel file is together with this html file in my Github repository, and the link to the government website is: 

https://www.energy.gov.au/publications/australian-energy-update-2019


I used the R code below, to transform the raw data into a format suitable for plotting in Tableau. The link to the resulting Tableau story is: 

https://public.tableau.com/views/AustraliasenergystorySeptember2021/Story


### Step 1
Import the source Excel sheet '2.2 Energy consumption by fuel type' as a data frame in R Studio.

```{r}
#Import the sheet into a variable called wide1
wide1 <- read_excel("Australian Energy Update 2020 data for charts.xlsx", 
    sheet = "2.2", range = "A4:E45")
wide1
```



### Step 2
Convert the imported data frame from wide to long format, then export it as a CSV file.
```{r}
#Use gather() to convert it to long format, and export this as a CSV file
long1 <- wide1 %>% gather(key = Fuel_type, value = Petajoules, `Coal`:`Renewables`)
write.csv(long1, "2.2 Energy consumption by fuel type.csv", row.names = FALSE)
long1
```

I repeated Steps 1 and 2 for each relevant Excel sheet, and shared the resulting insights in the 'Tableau story' above.



### End of R Markdown report
